
\documentclass[12pt]{article}
%	options include 12pt or 11pt or 10pt
%	classes include article, report, book, letter, thesis
\usepackage{tikz}
\usepackage{graphicx}

\title{Honors Calculus II: META}
\author{Brendan Whitaker}
\date{11 January 2016}

\begin{document}
\maketitle

\part{META}
\section{Curriculum Timeline}
\subsection{Chapters 15-17: Conics, Polars, and Parametrics}
This is the first third of the class and the material covered in this portion will substantiate the 1st Midterm. The date of the midterm is posted on the syllabus. 
\subsection{Vector Algebra \& A Little Bit of Differential Multivariable Calculus}
\subsection{Multivariable Calculus}
\section{Exams}
The first two thirds of the class will each have a corresponding Midterm exam. The Final will be a 2 hour exam. The first hour will be material from the last third of the class, Multivar. The second hour will be assorted topics which will be assigned at some point in the last third of the class. Exams will be given on one sheet of paper, blank loose leaf will be given for final solutions. 


\section{Homework and Classwork}
There will be 1 theoretical topic or problem each week. The rest of the material will be simple advanced calculation. The homework counts for the equivalent of 1 Exam within the grade scale. It will be collected each Friday, and the professor will attempt to give us a weekend to do each assignment when possible. The homework is not very heavy so it is imperative to do extra problems each week. There will be exactly 8 problems per assignment. Group work is encouraged. Calculators are to be used on the homework. NO CALCULATORS ON EXAMS. 
\section{General Guidelines}
Read the book!!!

The professor will cover highlights in class. You must read and know all of 15.1-15.5. 




 




\end{document}